---
title: 'Stats for CogSci2020 submission'
date: "May 2020"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, progress = TRUE)
```

```{r packageload, message=FALSE}
library(here)
library(tidyverse)
library(lme4)
library(RColorBrewer)
```


Read in the Zapotec dataframe with complete features.

```{r readdata, message=FALSE}
df <- read_csv(here("code", "data", "df_zapotec_fullfeats_cogsci2020.csv"))

````

Add variables we'll need for the analyses.

```{r preprocessing, message=FALSE}

slab <- function(z) {
  z[!(z %in% c("xlútsǐ", "càrpìntêr",  "msì", "bǎd",  "mtsùu", "wǐt", "dzǐn̲g"))] <-"other"
  z
}

df_mf <- df  %>% 
  mutate(is_named = !is.na(folk_specific), logmass = log(mass), logfreq = log(freq)) %>% 
  mutate(is_prot = !is.na(folk_specific) & folk_specific==folk_generic)  %>% 
  mutate(is_compound = grepl('-', folk_specific)) %>% 
  mutate(name_length = str_count(folk_specific) - is_compound) %>% 
  mutate(logname_length = log(name_length)) %>% 
  mutate(clements_group=factor(clements_group))  %>% 
  # XXX: include only species for which we have mass and frequency
  filter(mass > 0 & freq > 0) %>% 
  mutate(name=factor(folk_generic)) %>% 
  mutate(name=fct_relabel(name, slab)) %>% 
  mutate(name=factor(name, levels =  c("dzǐn̲g", "xlútsǐ", "càrpìntêr",  "msì", "bǎd",  "mtsùu", "wǐt", "other"))) 

  
# make long data frame for generating figures
df_mf_long <- df_mf %>% 
  gather(key='variable', value='val', logmass, logfreq)
```



## Named vs Unnamed categories

```{r namedunnamed, message=FALSE}

naming_full <- glm( formula = is_named ~ logfreq + logmass,  family = binomial, data = df_mf)
naming_freq<- glm( formula = is_named ~ logfreq,  family = binomial, data = df_mf)
naming_mass<- glm( formula = is_named ~ logmass, family = binomial, data = df_mf)
anova(naming_freq, naming_full, naming_mass, test="Chisq")


naming_drop <- drop1(naming_full, test="Chisq")
summary(naming_full)
naming_drop
```

The coefficients suggest that logfreq is a strong predictor (123.525) but that logmass is not (0.006). The call to `drop1` compares the full model against models that drop one of the predictors. Dropping logfreq means that the model fit gets significantly worse (p < 2e-16), but dropping logmass doesn't hurt the fit (p = 0.94). The AIC values support this conclusion (dropping logfreq makes the AIC bigger (ie worse), but dropping logmass makes the AIC a little bit better (smaller))


## Category size analysis
Now we analyze how many other species each named species is lumped with. 

```{r lumping, message=FALSE}
df_named <- df_mf %>% 
  filter(is_named) %>% 
  mutate(norm_group_size_generic = group_size_folk_generics/group_size_clements) 


lmlumping_full <- lm( formula = group_size_folk_generics ~ logfreq + logmass,  data = df_named)
lmlumping_freq <- lm( formula = group_size_folk_generics ~ logfreq,  data = df_named)
lmlumping_mass <- lm( formula = group_size_folk_generics ~ logmass,  data = df_named)

anova(lmlumping_freq, lmlumping_full, lmlumping_mass, test="Chisq")

summary(lmlumping_full)
```
The coefficients suggest that logmass (-1.18) is a stronger predictor than logfreq (-0.55). Dropping either predictor, however, leads to a model fit that is significantly worse (p values of 0.02 for logfreq and 2e-8 for logmass). AIC values support the same conclusion.


## Name length

```{r namelength, message=FALSE}


namelength_full <- lm( formula = logname_length ~ logfreq + logmass,  data = df_named)
namelength_freq <- lm( formula = logname_length ~ logfreq,  data = df_named)
namelength_mass <- lm( formula = logname_length ~ logmass,  data = df_named)

anova(namelength_freq, namelength_full, namelength_mass, test="Chisq")

summary(namelength_full)
```

Coefficients suggest that logfreq (0.04) and logmass (-0.06) are both weak predictors. Dropping logmass significantly impairs model performance (p = 0.03) but dropping logfreq doesn't (p = 0.16). AIC scores support same conclusion.


## Compound vs monomial


```{r compoundvsmonomial, message=FALSE}

logitmodcompound <- glm(
  formula = is_compound ~ logfreq + logmass, 
  family = binomial,
  data = df_named)

compound_drop <- drop1(logitmodcompound, test="Chisq")
compound_drop
summary(logitmodcompound)
```

Coefficients suggest that logfreq (-0.0387) is a stronger predictor than logmass (-0.0384) -- and the negative values indicating that if the frequency or mass of a species increase, it's *less* likely to have a compound name. But the `drop1` analyses suggest that model fit doesn't become significantly worse when either predictor is dropped (p values of 0.75 and 0.72). AIC scores support same conclusion.







